\documentclass[a4paper,11pt]{article}
\usepackage[lmargin=1.5cm,rmargin=1.5cm,bmargin=1.5cm]{geometry}
\usepackage[UKenglish]{babel}
\usepackage{graphicx}
\usepackage{courier}
\usepackage{color}
\usepackage{array}
\usepackage{pstricks}
\usepackage{parskip}
\usepackage[bottom]{footmisc}
\usepackage{fancyhdr}
\usepackage[none]{hyphenat}
\usepackage{soul}
% the hyperref package must always be the last package to be included
\usepackage[pdftex,
            pdfusetitle,
            pdfsubject={User's Manual for dastaZ80 homebrew computer},
            pdfkeywords={Z80, homebrew computer, Operating System, OS}
        ]{hyperref}

\hypersetup{
    colorlinks = true,
    linkcolor = blue,
    anchorcolor = blue,
    citecolor = blue,
    filecolor = blue,
    urlcolor = blue
}

\setlength\parindent{0pt}

\begin{document}
    \pagestyle{empty}
    % ==========================================================================
    % Cover page
    % ==========================================================================
    \begin{pspicture}(8.5,11)
        \rput[b](3.5,8){
            \parbox{10in}{
                \begin{flushright}
                    \Huge\bfseries\sffamily dastaZ80 Mark III\\ User's Manual
                \end{flushright}
            }
        }
        \uput[0](0,0){\color{blue}\rule{10in}{0.5ex}}
    \end{pspicture}
    \title{dastaZ80 User's Manual}
    \author{David Asta}
    \date{7 September 2022}

    \pagebreak
    % ==========================================================================
    % Header & Footer
    % ==========================================================================
    \pagestyle{fancy}
    \fancyhf{}
    \fancyhead[R]{dastaZ80 Mark III User's Manual}
    
    \begingroup
        \let\clearpage\relax
        \include{disclaimer_licenses}
        
        \textcopyright 2023 David Asta
    \endgroup

    \pagebreak
    % ==========================================================================
    \section*{Document Conventions}
    % ==========================================================================
    The following conventions are used in this manual:

    \begin{center}
        \begin{tabular}{c m{14cm}}
            \hline
            MUST & MUST denotes that the definition is and absolute
            requirement.\\
            \hline
            SHOULD & SHOULD denotes that it is recommended, but that there may
            exist valid reasons to ignore it.\\
            \hline
            \textbf{DEVICE} & Device names are displayed in bold all upper case 
            letters, and refer to hardware devices.\\
            \hline
            \textbf{command} & Operating System command keywords are displayed 
            in bold all lower case letters.\\
            \hline
            \textit{$<$text$>$} & Angle brackets enclose variable information 
            that you MUST supply. In place of \textit{$<$text$>$}, substitute 
            the value desired. Do not enter the angle brackets when entering the
            value.\\
            \hline
            \textit{$[$text$]$} & Square brackets enclose variable information
            that you COULD supply. They are optional. In place of $[$text$]$,
            substitute the value desired. Do not enter the square brackets when
            entering the value.\\
            \hline
            \texttt{Courier} & Text appearing in the \texttt{Courier} font 
            represents information that you type in via the keyboard.\\
            \hline
            0x14B0 & Numbers prefixed by 0x indicate an Hexadecimal value.
            Unless specified, memory addresses are always expressed in
            Hexadecimal.\\
            \hline
            \textit{Return} & Refers to the key Return in the keyboard.\\
            \hline
        \end{tabular}
    \end{center}

    \hfill\break
    \hfill\break

    The SD card is referred as \textbf{DISK}.

    The Floppy Disk Drive is referred as \textbf{DISK} or as \textbf{FDD}.

    The 80 column text VGA output is referred as \textbf{CONSOLE} or as
    \textbf{High Resolution Display}.

    The 40 column graphics Composite Video output is referred as \textbf{Low
    Resolution Display}.
    
    The Operating System may be referred as DZOS, dzOS or simply OS.

    \textbf{MEMORY} refers to both \textbf{ROM} and \textbf{RAM}.

    Although the word \textbf{ROM} is used in this manual, the actual chip used
    in the dastaZ80 computer is of the type \textbf{EEPROM}.

    Memory used by the \textbf{Low Resolution Display} is referred as
    \textbf{VRAM} (Video RAM).

    \pagebreak
    % ==========================================================================
    \section*{Related Documentation}
    % ==========================================================================
    \begin{itemize}
        \item dastaZ80 User's Manual\cite{dastaz80userman}
        \item dastaZ80 Technical Reference Manual\cite{dastaz80techman}
        \item dzOS Github Repository\cite{dastaZ80github}
        \item Software for dzOS Github Repository\cite{dastaZ80githubsoft}
        \item Nascom 2 Microcomputer BASIC Programming Manuals\cite{nascombasic}
        \item Z80 Family CPU User Manual\cite{z80manual}
    \end{itemize}

    \pagebreak
    % ==========================================================================
    \tableofcontents
    % ==========================================================================

    \pagebreak
    % ==========================================================================
    % Header & Footer
    % ==========================================================================
    \pagestyle{fancy}
    \fancyhf{}
    \fancyhead[R]{dastaZ80 Mark III User's Manual}
    \fancyfoot[R]{\thepage}
    \setcounter{page}{1}
    % ==========================================================================
    \section{Introduction}
    % ==========================================================================
    The dastaZ80 is a homebrew computer designed and built following the style
    of the 8-bit computers of the 80s that I used on those days: Amstrad CPC,
    Commodore 64 and MSX. The name comes from “d”avid “asta” (my name) and 
    “Z80” (the CPU used).

    The idea behind the making of this computer came from an initial wish of
    writing an operating system (OS) for an 8-bit machine. Not comfortable with
    writing an OS for an already existing computer like an Amstrad CPC, C64,
    MSX, etc., due to the complexity of its hardware (or rather my lack of
    knowledge), I decided to built my own 8-bit computer from scratch, so that
    I could fully understand the hardware and also influence the design.

    The OS written by me for this computer is called \textit{DZOS}, from 
    dastaZ80 OS. Sometimes I spell it as \textit{dzOS}. Haven’t made my mind 
    yet.

    This manual describes the usage of DZOS running on the dastaZ80 computer.
    
    \pagebreak
    % ==========================================================================
    \section{dastaZ80 Overview}
    % ==========================================================================

    The dastaZ80 computer comes in two different formats:

    \begin{itemize}
        \item dastaZ80 (Original), in Acorn Archimedes A3010 case.
        \item dastaZ80DB (\textit{DB} stands for Desktop Box), in a desktop
            computer box.
    \end{itemize}

    Both are functionally the same, except the dastaZ80DB does not have a
    keyboard nor a Floppy Disk Drive, and is meant to be operated from an
    external video terminal (like DEC VT100) or a computer running a terminal
    program (like Minicom or PuTTY). More details about this configuration in
    the section \hyperref[sec:setting_system]{Setting up the system}.

    The dastaZ80 Original was the first version I made, and it is a stand-alone
    computer. Only things needed to use it are: the computer itself, a 5V/4A
    power supply, a monitor with VGA connector and optionally a monitor with 
    NTSC Composite video input and (also optionally) a pair of amplified
    speakers or a monitor with RCA stereo connectors.

    The dastaZ80DB requires the same, plus an external serial keyboard. Also,
    the VGA can be exchanged by a serial monitor.

    % ==========================================================================
    \subsection{dastaZ80 Original}
    % ==========================================================================

    \begin{itemize}
        \item ZiLOG Z80 microprocessor (CPU) running at 7.3728 MHz
        \item 64 KB RAM: 16 KB reserved for DZOS, 48 KB available
            for the user and programms.
        \item Storage devices: Micro SD Card\footnote{Formatted with FAT32 and
            containing Disk Image Files formatted with DZFS (dastaZ80 File
            System.)}, 3.5" DD/HD Floppy Disk Drive.
        \item Dual Video output: VGA 80 columns by 25 lines and 16 colours,
            Composite NTSC 15 colours.
        \item Stereo Sound: 3 channels (1 tone per channel), 7 octaves, 1 noise
            channel, envelope generator\footnote{Same chip (AY-3-8912 PSG) as
            used in numerous Arcade machines and in Amstrad CPC, Atari ST, MSX,
            ZX Spectrum and other home computers.}.
        \item Real-Time Clock: Date and Time backed up with button cell battery.
        \item Keyboard: Acorn Archimedes A3010 keyboard. 102 keys with 12 function
            keys, cursor keys and numeric pad.
        \item Case: Repurposed Acorn Archimedes A3010 case with keyboard.
        \item Expansion ports: GPIO and ROM Cartridges.
    \end{itemize}

    % ==========================================================================
    \subsection{dastaZ80DB}
    % ==========================================================================

    Same as above, with the only differences:

    \begin{itemize}
        \item No keyboard.
        \item No Floppy Disk Drive.
        \item Case: desktop box.
    \end{itemize}

    \pagebreak
    \include{sections/settingup}
    \pagebreak
    \include{sections/operation}
    \pagebreak
    \include{sections/maintenance}
    \pagebreak
    \include{sections/os}
    \pagebreak
    \include{sections/os_commands}
    \pagebreak
    \include{sections/other_software}
    \include{sections/appendixes}
    \pagebreak
    \include{sections/glossary}
    \pagebreak

    % ==========================================================================
    % Bibliography
    % ==========================================================================
    \pagebreak
    \bibliographystyle{unsrt}
    \bibliography{../dastaZ80}
\end{document}
